Quantifying social information in natural infant visual experience

### 2020-01-12 analysis/analysis_fulldataset.Rmd
more intro

### Load file to filter flipped videos (when sam's mom was wearing camera; not egocentric)
```{r}
right_side_up_file = "../data/video_right-side-up.csv"
right_side_up =read.csv(right_side_up_file)

@@

### Load SAYCAM metadata from csv file; get metadata from detections .csv
```{r}
metadata_file = "../data/SAYcam-2019-12-0922_54_28.csv"

>>>>

### Load file to filter flipped videos (when sam's mom was wearing camera; not egocentric)
```{r}
right_side_up_file = "../src/data/video_right-side-up.csv"
right_side_up =read.csv(right_side_up_file)

@@

### Load SAYCAM metadata from csv file; get metadata from detections .csv
```{r}
metadata_file = "../src/data/SAYcam-2019-12-0922_54_28.csv"

### 2020-01-12 paper/cogsci/paper.Rmd
more intro

abstract: >
    Faces are a critical part of infants' visual experience, as they convey social and 
    linguistic information.

keywords: >
    Add your choice of indexing terms or keywords; kindly use a semi-colon; between each term.

@@

Previous work has suggested:
- Relative changes in prevalence of faces vs. hands (Fausey 2016), though perhaps this is driven by infants younger than 4 months of age (e.g., Jayaraman, Fausey, & Smith, 2015; Sugden, Mohamed-Ali, & Moulson, 2014) who see both more frequent and more persistent faces (Jayaraman & Smith, 2018)

- Surprising attention to hand movements and their interactions with objects (Yu & Smith, 2013), particularly in older infants

- Differences in the availability of social information depending on motor abilities (Sanchez, Long et al., 2018 CogSci; @Franchak2011 and more)

@@

Here, we analyze a longitudinal corpus of head-mounted camera data, leveraging over XX hours of videos from XX children, totaling more than XX frames.  
(briefly describe dataset; sampling strategy: location of two households, number of hours of video, variability in location, etc; reference published paper on what this dataset is, large field of view (fisheye lens))

@@

Part 1: How well can we capture social information using computer vision?
- Description of OpenPose (Figure 1) 
- Description of annotation strategy (24K by Ketan, ~4K on turk, reliability)
- Describe main P/R/F statistics for 24K for faces and hands; interpret. Relatively higher precision vs. recall.
- P/R/F variation across child/age for faces
- Describe possible sources of variation that decrease scores for:
    - Faces: weird viewpoints, occluded/side viewpoint, faces in books
    - Hands: children’s own hands, hands in books, side viewpoints
- Describe additional child vs. hand annotation; P/R/F variation across child vs. adult hands (better for adult hands, still OK for child hands)

Part 2: Access to social information across age
Prevalence of hands vs faces across age (in goldset, full dataset) (Figure 2)
Why so many hands? 
More child hands as in gold set
Cropped analysis

Variation across location contexts (goldset, full dataset)

@@

If the authors' names are
included in the sentence, place only the year in parentheses, as in
[-@Franchak2011], but otherwise place the entire reference in
parentheses with the authors and year separated by a comma
[@Fausey2016]. List multiple references alphabetically and
separate them by semicolons [@Frank2012; @Smith2011]. 
Use the et. al. construction only after listing all the authors to a
publication in an earlier reference and for citations with four or
more authors.

>>>>

abstract: >
    The faces and hands of infants' caregivers and other social partners offer a rich source of social and
    causal information that may be critical for infants' cognitive and linguistic development.

keywords: >
    social cognition; face perception; infancy; head cameras; deep learning

@@

Infants are famously confronted by a blooming, buzzing onslaught of stimuli (James, 1891) which they must learn to parse and navigate as their cognitive and social skills develop.
Fortunately, they can depend on regularities not only in their visual environments [@Aslin2009] and linguistic environments, but also in the presence and actions of their caregivers and other social partners [@].
From a very young age, infants show great attention to faces CITE, and indeed faces are a critical part of infants' visual experience as an important conduit of social and linguistic information.
As infants mature, they begin to also attend more to the manual actions taken by their social partners, engaging in joint attention to objects and events around them. 
Such episodes are prompted not only by a glance from a caregiver, but also through pointing and offering; thus, hands are also an important carrier of information, especially relevant for learning language and actions.
It is unsurprising, then, that both hands and faces are prevalent in even young infants' visual experiences, as evidenced through analyses of egocentric views collected with head-mounted cameras (i.e., headcams; CITES).

Of course, not only is children's interest and attention to hands and faces

Previous work has suggested:

Previous work has found changes in the prevalence of faces vs. hands for infants of different ages. For example, Fausey et al. [-@Fausey2016] found that infants less than 12 months of age received face-dense input, relative to 1- to 2-year-olds who received more hand-dense input. 
However, it may be that this effect is driven by infants younger than 4 months of age [e.g., @Jayaraman2015; @Sugden2014] who see both more frequent and more persistent faces [@Jayaraman2018].
Once infants begin to crawl (~6 months) they may see far fewer hands and faces, overall.

Earlier work has also found surprising attention to hand movements and their interactions with objects (Yu & Smith, 2013), particularly in older infants (ages?).

Differences in the availability of social information depending on motor abilities [@Sanchez2018; @Franchak2011] (more Franchak papers?)

@@

Here, we analyze the SAYcam dataset [@SAYcam], a longitudinal corpus of head-mounted camera data comprising more than 1700 videos from three children, for a total of over 300 hours of videos (>100 million frames). 
Over a span of 6 to 32 months of age, the three children (S, A, and Y) in the dataset wore headcams at least twice weekly, for approximately one hour per recording session. 
One weekly session was on the same day each week at a roughly constant time of day, while the other(s) were chosen arbitrarily at the participating family's discretion.

This dataset differs in four key ways: 1) number of hours/frames 2) not just mealtimes--naturalistic sample of many contexts, 3) longitudinal, 4) much larger field-of-view.

@@

@@

## Dataset

(briefly describe dataset; sampling strategy: location of two households, number of hours of video, variability in location, etc; reference published paper on what this dataset is, large field of view (fisheye lens))
(109 degrees horizontal x 70 degrees vertical)

## Part 1: How well can we capture social information using computer vision?

### Description of OpenPose (Figure 1) 
To automatically annotate the millions of frames in SAYcam, we use OpenPose [@Cao2018openpose; @Simon2017hand], a computer vision model optimized for jointly detecting human face, body, hand, and foot keypoints (135 in total) that operates well on scenes including multiple people even if they are partially-occluded.

### Description of annotation strategy (24K by Ketan, 4K on Amazon Mechanical Turk, reliability)
To test the validity of OpenPose's hand and face detections, we compared to human annotations of 24,000 frames selected uniformly at random from the videos of two children (S and A).

### Describe main PRF statistics for 24K for faces and hands; interpret.
Relatively higher precision vs. recall.
- P/R/F variation across child/age for faces
- Describe possible sources of variation that decrease scores for:
    - Faces: weird viewpoints, occluded/side viewpoint, faces in books
    - Hands: children's own hands, hands in books, side viewpoints
- Describe additional child vs. hand annotation; P/R/F variation across child vs. adult hands (better for adult hands, still OK for child hands)

## Part 2: Access to social information across age

### Prevalence of hands vs faces across age (in goldset, full dataset) (Figure 2)

### Why so many hands?

### More child hands as in gold set

## Field-of-View Comparison

The field of view (FOV) of the fisheye lens used in @SAYcam is much wider (109 degrees horizontal x 70 degrees vertical) than the FOV of the lens used in @Fausey2016 (69 deg. x 41 deg.). 
Looks like child hands make up about ~34% of the hands detected in our gold set (in Fausey 2016, they are only 8% of the hands). 
Furthermore, a lot of the lower proportions of hands come from the infants <6 months of age.

## Variability by Location

Next we examine variation in the presence of hands and faces across different locations.
Of the 3,027 videos, the content of 1,829 have been manually manuallly annotated for filming location, activities taking place, and visible objects (see @SAYcam). 
To give a sense of the contexts the children experienced, the most frequent filming locations were the living room (339 videos), bedroom (182), kitchen (150), outside on property (129), child's bedroom (81), deck/porch (73), hallway (70), and off property (57). Filming only took place twice in the dining room.

The most frequent activities were sitting (410), playing (375), being held (352), and standing (297). Eating was the 11th most-frequent activity (117 videos).

(goldset, full dataset)

List multiple references alphabetically and
separate them by semicolons [@Frank2012; @Smith2011].

### 2020-01-12 paper/cogsci/paper.tex
more intro

\begin{abstract}
Faces are a critical part of infants' visual experience, as they convey
social and linguistic information.

\textbf{Keywords:}
Add your choice of indexing terms or keywords; kindly use a semi-colon;
between each term.
\end{abstract}

@@

Previous work has suggested: - Relative changes in prevalence of faces
vs.~hands (Fausey 2016), though perhaps this is driven by infants
younger than 4 months of age (e.g., Jayaraman, Fausey, \& Smith, 2015;
Sugden, Mohamed-Ali, \& Moulson, 2014) who see both more frequent and
more persistent faces (Jayaraman \& Smith, 2018)

\begin{itemize}
\item
  Surprising attention to hand movements and their interactions with
  objects (Yu \& Smith, 2013), particularly in older infants
\item
  Differences in the availability of social information depending on
  motor abilities (Sanchez, Long et al., 2018 CogSci; Franchack papers)
\end{itemize}

@@

Here, we analyze a longitudinal corpus of head-mounted camera data,
leveraging over XX hours of videos from XX children, totaling more than
XX frames. \{briefly describe dataset; sampling strategy: location of
two households, number of hours of video, variability in location, etc;
reference published paper on what this dataset is, large field of view
(fisheye lens))

To do so, we first test and validate novel computer vision methods for
extracting social information from these egocentric viewpoints on a
small subset of randomly selected frames from the dataset. We then apply
these methods at scale to the larger dataset, allowing us to extract key
descriptive variables hypothesized to vary across development.

Part 1: How well can we capture social information using computer
vision? - Description of OpenPose (Figure 1) - Description of annotation
strategy (24K by Ketan, \textasciitilde{}4K on turk, reliability) -
Describe main P/R/F statistics for 24K for faces and hands; interpret.
Relatively higher precision vs.~recall. - P/R/F variation across
child/age for faces - Describe possible sources of variation that
decrease scores for: - Faces: weird viewpoints, occluded/side viewpoint,
faces in books - Hands: children's own hands, hands in books, side
viewpoints - Describe additional child vs.~hand annotation; P/R/F
variation across child vs.~adult hands (better for adult hands, still OK
for child hands)

Part 2: Access to social information across age Prevalence of hands vs
faces across age (in goldset, full dataset) (Figure 2) Why so many
hands? More child hands as in gold set Cropped analysis

Variation across location contexts (goldset, full dataset)

@@

If the authors' names are included in the sentence, place only the year
in parentheses, as in (2011), but otherwise place the entire reference
in parentheses with the authors and year separated by a comma (Fausey,
Jayaraman, \& Smith, 2016). List multiple references alphabetically and
separate them by semicolons (Frank, 2012; Smith, Yu, \& Pereira, 2011).
Use the et. al.~construction only after listing all the authors to a
publication in an earlier reference and for citations with four or more
authors.

@@

\begin{table}[H]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & t value & Pr($>$$|$t$|$) \\ 
  \hline
(Intercept) & 0.00 & 0.10 & 0.0 & 0.98 \\ 
  x & 1.90 & 0.09 & 20.1 & 0.00 \\ 
   \hline
\end{tabular}
\caption{This table prints across one column.} 
\end{table}

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Fausey2016}{}%
Fausey, C. M., Jayaraman, S., \& Smith, L. B. (2016). From faces to
hands: Changing visual input in the first two years. \emph{Cognition},
\emph{152}, 101--107.

@@

@@

>>>>

\begin{abstract}
The faces and hands of infants' caregivers and other social partners
offer a rich source of social and causal information that may be
critical for infants' cognitive and linguistic development.

\textbf{Keywords:}
social cognition; face perception; infancy; head cameras; deep learning
\end{abstract}

@@

Infants are famously confronted by a blooming, buzzing onslaught of
stimuli (James, 1891) which they must learn to parse and navigate as
their cognitive and social skills develop. Fortunately, they can depend
on regularities not only in their visual environments (Aslin, 2009) and
linguistic environments, but also in the presence and actions of their
caregivers and other social partners {[}@{]}. From a very young age,
infants show great attention to faces CITE, and indeed faces are a
critical part of infants' visual experience as an important conduit of
social and linguistic information. As infants mature, they begin to also
attend more to the manual actions taken by their social partners,
engaging in joint attention to objects and events around them. Such
episodes are prompted not only by a glance from a caregiver, but also
through pointing and offering; thus, hands are also an important carrier
of information, especially relevant for learning language and actions.
It is unsurprising, then, that both hands and faces are prevalent in
even young infants' visual experiences, as evidenced through analyses of
egocentric views collected with head-mounted cameras (i.e., headcams;
CITES).

Of course, not only is children's interest and attention to hands and
faces

Previous work has suggested:

Previous work has found changes in the prevalence of faces vs.~hands for
infants of different ages. For example, Fausey et al. (2016) found that
infants less than 12 months of age received face-dense input, relative
to 1- to 2-year-olds who received more hand-dense input. However, it may
be that this effect is driven by infants younger than 4 months of age
(e.g., Jayaraman, Fausey, \& Smith, 2015; Sugden, Mohamed-Ali, \&
Moulson, n.d.) who see both more frequent and more persistent faces
(Jayaraman \& Smith, 2018). Once infants begin to crawl
(\textasciitilde{}6 months) they may see far fewer hands and faces,
overall.

Earlier work has also found surprising attention to hand movements and
their interactions with objects (Yu \& Smith, 2013), particularly in
older infants (ages?).

Differences in the availability of social information depending on motor
abilities (Franchak, Kretch, Soska, \& Adolph, 2011; Sanchez, Long,
Kraus, \& Frank, 2018) (more Franchak papers?)

@@

Here, we analyze the SAYcam dataset (Sullivan, Mei, Perfors, Wojcik, \&
Frank, n.d.), a longitudinal corpus of head-mounted camera data
comprising more than 1700 videos from three children, for a total of
over 300 hours of videos (\textgreater{}100 million frames). Over a span
of 6 to 32 months of age, the three children (S, A, and Y) in the
dataset wore headcams at least twice weekly, for approximately one hour
per recording session. One weekly session was on the same day each week
at a roughly constant time of day, while the other(s) were chosen
arbitrarily at the participating family's discretion.

This dataset differs in four key ways: 1) number of hours/frames 2) not
just mealtimes--naturalistic sample of many contexts, 3) longitudinal,
4) much larger field-of-view.

To do so, we first test and validate novel computer vision methods for
extracting social information from these egocentric viewpoints on a
small subset of randomly selected frames from the dataset.\\
We then apply these methods at scale to the larger dataset, allowing us
to extract key descriptive variables hypothesized to vary across
development.

@@

\hypertarget{dataset}{%
\subsection{Dataset}\label{dataset}}

(briefly describe dataset; sampling strategy: location of two
households, number of hours of video, variability in location, etc;
reference published paper on what this dataset is, large field of view
(fisheye lens)) (109 degrees horizontal x 70 degrees vertical)

\hypertarget{part-1-how-well-can-we-capture-social-information-using-computer-vision}{%
\subsection{Part 1: How well can we capture social information using
computer
vision?}\label{part-1-how-well-can-we-capture-social-information-using-computer-vision}}

\hypertarget{description-of-openpose-figure-1}{%
\subsubsection{Description of OpenPose (Figure
1)}\label{description-of-openpose-figure-1}}

To automatically annotate the millions of frames in SAYcam, we use
OpenPose (Cao, Hidalgo, Simon, Wei, \& Sheikh, 2018; Simon, Joo,
Matthews, \& Sheikh, 2017), a computer vision model optimized for
jointly detecting human face, body, hand, and foot keypoints (135 in
total) that operates well on scenes including multiple people even if
they are partially-occluded.

\hypertarget{description-of-annotation-strategy-24k-by-ketan-4k-on-amazon-mechanical-turk-reliability}{%
\subsubsection{Description of annotation strategy (24K by Ketan, 4K on
Amazon Mechanical Turk,
reliability)}\label{description-of-annotation-strategy-24k-by-ketan-4k-on-amazon-mechanical-turk-reliability}}

To test the validity of OpenPose's hand and face detections, we compared
to human annotations of 24,000 frames selected uniformly at random from
the videos of two children (S and A).

\hypertarget{describe-main-prf-statistics-for-24k-for-faces-and-hands-interpret.}{%
\subsubsection{Describe main PRF statistics for 24K for faces and hands;
interpret.}\label{describe-main-prf-statistics-for-24k-for-faces-and-hands-interpret.}}

Relatively higher precision vs.~recall. - P/R/F variation across
child/age for faces - Describe possible sources of variation that
decrease scores for: - Faces: weird viewpoints, occluded/side viewpoint,
faces in books - Hands: children's own hands, hands in books, side
viewpoints - Describe additional child vs.~hand annotation; P/R/F
variation across child vs.~adult hands (better for adult hands, still OK
for child hands)

\hypertarget{part-2-access-to-social-information-across-age}{%
\subsection{Part 2: Access to social information across
age}\label{part-2-access-to-social-information-across-age}}

\hypertarget{prevalence-of-hands-vs-faces-across-age-in-goldset-full-dataset-figure-2}{%
\subsubsection{Prevalence of hands vs faces across age (in goldset, full
dataset) (Figure
2)}\label{prevalence-of-hands-vs-faces-across-age-in-goldset-full-dataset-figure-2}}

\hypertarget{why-so-many-hands}{%
\subsubsection{Why so many hands?}\label{why-so-many-hands}}

\hypertarget{more-child-hands-as-in-gold-set}{%
\subsubsection{More child hands as in gold
set}\label{more-child-hands-as-in-gold-set}}

\hypertarget{field-of-view-comparison}{%
\subsection{Field-of-View Comparison}\label{field-of-view-comparison}}

The field of view (FOV) of the fisheye lens used in Sullivan et al.
(n.d.) is much wider (109 degrees horizontal x 70 degrees vertical) than
the FOV of the lens used in Fausey et al. (2016) (69 deg. x 41 deg.).
Looks like child hands make up about \textasciitilde{}34\% of the hands
detected in our gold set (in Fausey 2016, they are only 8\% of the
hands). Furthermore, a lot of the lower proportions of hands come from
the infants \textless{}6 months of age.

\hypertarget{variability-by-location}{%
\subsection{Variability by Location}\label{variability-by-location}}

Next we examine variation in the presence of hands and faces across
different locations. Of the 3,027 videos, the content of 1,829 have been
manually manuallly annotated for filming location, activities taking
place, and visible objects (see Sullivan et al. (n.d.)). To give a sense
of the contexts the children experienced, the most frequent filming
locations were the living room (339 videos), bedroom (182), kitchen
(150), outside on property (129), child's bedroom (81), deck/porch (73),
hallway (70), and off property (57). Filming only took place twice in
the dining room.

The most frequent activities were sitting (410), playing (375), being
held (352), and standing (297). Eating was the 11th most-frequent
activity (117 videos).

(goldset, full dataset)

List multiple references alphabetically and separate them by semicolons
(Frank, 2012; Smith, Yu, \& Pereira, 2011).

@@

\begin{table}[H]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & t value & Pr($>$$|$t$|$) \\ 
  \hline
(Intercept) & -0.09 & 0.10 & -0.9 & 0.35 \\ 
  x & 2.06 & 0.11 & 18.5 & 0.00 \\ 
   \hline
\end{tabular}
\caption{This table prints across one column.} 
\end{table}

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Aslin2009}{}%
Aslin, R. N. (2009). How infants view natural scenes gathered from a
head-mounted camera. \emph{Optometry and Vision Science: Official
Publication of the American Academy of Optometry}, \emph{86}(6),
561--565.

\leavevmode\hypertarget{ref-Cao2018openpose}{}%
Cao, Z., Hidalgo, G., Simon, T., Wei, S.-E., \& Sheikh, Y. (2018).
OpenPose: Realtime multi-person 2D pose estimation using Part Affinity
Fields. In \emph{ArXiv preprint arXiv:1812.08008}.

\leavevmode\hypertarget{ref-Fausey2016}{}%
Fausey, C. M., Jayaraman, S., \& Smith, L. B. (2016). From faces to
hands: Changing visual input in the first two years. \emph{Cognition},
\emph{152}, 101--107.

@@

\leavevmode\hypertarget{ref-Jayaraman2015}{}%
Jayaraman, S., Fausey, C. M., \& Smith, L. B. (2015). The faces in
infant-perspective scenes change over the first year of life. \emph{PLoS
One}. \url{http://doi.org/10.1371/journal.pone.0123780}

\leavevmode\hypertarget{ref-Jayaraman2018}{}%
Jayaraman, S., \& Smith, L. B. (2018). Faces in early visual
environments are persistent not just frequent. \emph{Vision Research}.

\leavevmode\hypertarget{ref-Sanchez2018}{}%
Sanchez, A., Long, B., Kraus, A. M., \& Frank, M. C. (2018). Postural
developments modulate children's visual access to social information. In
\emph{Proceedings of the 40th annual conference of the cognitive science
society}.

\leavevmode\hypertarget{ref-Simon2017hand}{}%
Simon, T., Joo, H., Matthews, I., \& Sheikh, Y. (2017). Hand keypoint
detection in single images using multiview bootstrapping. In
\emph{CVPR}.

@@

\leavevmode\hypertarget{ref-Sugden2014}{}%
Sugden, N. A., Mohamed-Ali, M. I., \& Moulson, M. C. (n.d.). I spy with
my little eye: Typical, daily exposure to faces documented from a
first-person infant perspective. \emph{Developmental Psychobiology},
\emph{56}(2), 249--261.

\leavevmode\hypertarget{ref-SAYcam}{}%
Sullivan, J., Mei, M., Perfors, A., Wojcik, E., \& Frank, M. (n.d.).
Head cameras on children aged 6 months through 31 months.
